---
title: "ST662 Projcet"
author: "Stephen Foley"
date: "3/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## packages
```{r echo = F}
library(tidyverse)
library(ggplot2)
library(highcharter)
library()
library()
library()
library()
library()
library()
library()
library()
library()

```

## reading in the data
```{r}
complaints <- read.csv('complaints1.csv')
str(complaints)

```



# Data cleaning

## merging duplicate credit reporting columns, filling in missing values for 'Credit reporting' sub.product
```{r}
complaints$Sub.product[complaints$Product == 'Credit reporting'] <- 'Credit reporting'

complaints$Product[complaints$Product == 'Credit reporting'] <- 'Credit reporting, credit repair services, or other personal consumer reports'


```

## merging duplicate money transfer columns
```{r}



complaints$Product[complaints$Product == 'Money transfers'] <- 'Money transfer, virtual currency, or money service'	

```


## merging duplicate credit card columns, filling in missing values for 'Credit card' sub.product with cerd card
```{r}

complaints$Sub.product[complaints$Product == 'Credit card'] <- 'Credit card'

complaints$Product[complaints$Product == 'Credit card'] <- 'Credit card or prepaid card'



```


## merging duplicate Payday loan columns, filling in missing values
```{r}
complaints$Sub.product[complaints$Product == 'Payday loan'] <- 'Payday loan'

complaints$Product[complaints$Product == 'Payday loan'] <- 'Payday loan, title loan, or personal loan'


```






## recoding Company response to consumer column as either “Closed with monetary relief” or “Other”
```{r}

complaints$Company.response.to.consumer[complaints$Company.response.to.consumer != 'Closed with monetary relief'] <- 'Other'


```


## to replace missing values with NA
```{r}
complaints[complaints == ''] <- NA

```


































## to see which companies have the most complaints 
```{r}
top_com <- complaints %>% select(Company) %>%
  group_by(Company) %>%
  summarise(count = n()) %>% arrange(-count)

```


## to see what products have the most complaints
```{r}
table(complaints$Product)

top_prod <- complaints %>% select(Product) %>%
  group_by(Product) %>% 
  summarise(count = n())

AM <- top_prod %>% 
  hchart('column',hcaes(x='Product',y='count', group='Product'),pointwidth=3)
AM <- AM %>% hc_title(text = "Most common product complaints",
                      style = list(fontWeight = "bold"),
                      align = "center") %>%
  hc_xAxis(title = list(text = ""),
           style = list(fontWeight = "bold")) %>%
  hc_yAxis(title = list(text = "complaints"),
           style = list(fontWeight = "bold")) %>%
  hc_legend(enabled = F)
AM
```

## time series of complaints for top 10 companies (prep code)
```{r}
top_com_names <- top_com$Company[1:10]
comt <- complaints %>% select(Date.received, Company) %>%
  filter(str_detect(Company, top_com_names))



comt$Date.received <- gsub("[[:punct:]]", "",comt$Date.received)
comt$Date.received <- as.Date(comt$Date.received,"%Y%m%d")

comt1 <- comt
comt1$month <- strftime(as.POSIXct(comt$Date.received), "%m")
comt1$year=strftime(as.POSIXct(comt$Date.received), "%Y")


comt2 <- comt1
comt2$Date <- as.POSIXct(paste(comt2$year, comt2$month, "01", sep = "-"),'GMT')
comt2 <- comt2 %>% select(Date, Company) %>% 
  group_by(Date, Company) %>%
  summarise(count = n())
comt2$Date <- as.Date(comt2$Date)
comt2$Company <- as.factor(comt2$Company)




```

## time series of complaints for top 10 companies (plot code)
```{r}
A <- comt2 %>% hchart('line',hcaes(x=Date, y=count, group=Company))
A <- A %>% 
  hc_title(text = "Time series of complaints per company",
           style = list(fontWeight = "bold"),
           align = "center") %>%
  hc_xAxis(title = list(text = "Date"),
           style = list(fontWeight = "bold")) %>%
  hc_yAxis(title = list(text = "Count"),
           style = list(fontWeight = "bold")) %>%
  hc_legend(title = list(text= 'Company'),layout = "vertical", verticalAlign = 'middle',
            align = "right") %>%
  hc_tooltip(crosshairs = TRUE,
             shared = TRUE, 
             borderWidth = 4) %>% 
  hc_exporting(enabled = TRUE)
A


```




## time series of product complaints for all companies
```{r}
prodt <- complaints %>% select(Date.received, Product) 



prodt$Date.received <- gsub("[[:punct:]]", "",prodt$Date.received)
prodt$Date.received <- as.Date(prodt$Date.received,"%Y%m%d")

prodt1 <- prodt
prodt1$month <- strftime(as.POSIXct(prodt1$Date.received), "%m")
prodt1$year=strftime(as.POSIXct(prodt1$Date.received), "%Y")


prodt2 <- prodt1
prodt2$Date <- as.POSIXct(paste(prodt2$year, prodt2$month, "01", sep = "-"),'GMT')
prodt2 <- prodt2 %>% select(Date, Product) %>% 
  group_by(Date, Product) %>%
  summarise(count = n())
prodt2$Date <- as.Date(prodt2$Date)
prodt2$Product <- as.factor(prodt2$Product)


A <- prodt2 %>% hchart('line',hcaes(x=Date, y=count, group=Product))
A <- A %>% 
  hc_title(text = "Time series of product complaints by type",
           style = list(fontWeight = "bold"),
           align = "center") %>%
  hc_xAxis(title = list(text = "Date"),
           style = list(fontWeight = "bold")) %>%
  hc_yAxis(title = list(text = "Count"),
           style = list(fontWeight = "bold")) %>%
  hc_legend(enabled = T) %>%
  hc_tooltip(crosshairs = TRUE,
             shared = TRUE, 
             borderWidth = 4) %>% 
  hc_exporting(enabled = TRUE)
A


```

## to see which products are most commonly closed with monetary relief
```{r}

mont <- complaints %>% select( Product, Company.response.to.consumer)
mont <- mont %>% group_by(Product, Company.response.to.consumer) %>%
  summarise(count = n())
mont$Product <- as.factor(mont$Product)
mont$Company.response.to.consumer <- as.factor(mont$Company.response.to.consumer)


mont1 <- mont %>% pivot_wider(names_from = Company.response.to.consumer, values_from = count)

mont1 <- mont1 %>% rename(
 Closed_with_monetary_relief =  `Closed with monetary relief` 
)
mont1 <- mont1 %>% group_by(Product) %>%
  mutate(
    prop = 100*(Closed_with_monetary_relief/( Closed_with_monetary_relief + Other ))
  )

AM <- mont1 %>% 
  hchart('column',hcaes(x='Product',y='prop', group = 'Product'),pointwidth=5)
AM <- AM %>% hc_title(text = "Proportion of products closed with monetary relief",
                      style = list(fontWeight = "bold"),
                      align = "center") %>%
  hc_xAxis(title = list(text = ""),
           style = list(fontWeight = "bold")) %>%
  hc_yAxis(title = list(text = "Percentage %"),
           style = list(fontWeight = "bold")) %>%
  hc_legend(enabled = F)
AM

```



## stacked barplot of product complaint categories
```{r}

ice <- complaints %>% select(Date.received, Product)
ice$Date.received <- gsub("[[:punct:]]", "",ice$Date.received)
ice$Date.received <- as.Date(ice$Date.received,"%Y%m%d")


ice1 <- ice
ice1$year=strftime(as.POSIXct(ice1$Date.received), "%Y")
str(ice1)


ice1$Date <- as.POSIXct(paste(ice1$year, "01", "01", sep = "-"),'GMT')
ice1 <- ice1 %>% select(Date, Product) %>% 
  group_by(Date, Product) %>%
  summarise(count = n())
ice1$year <- as.Date(ice1$year)
ice1$Product <- as.factor(ice1$Product)

ggplot(ice1, aes(fill = Product, y=count, x= Date)) + 
  geom_bar(position = 'fill', stat = 'identity') + 
  scale_fill_viridis_d()


```